
# Categorical data (facotr)

```{r, echo = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```
## Frequency and contingency table



###  `frq()` and `flat_table()`   


sjmisc

https://strengejacke.github.io/sjmisc/articles/exploringdatasets.html  

```{r}
library(sjmisc)
```

```{r}
data(efc)
(efc <- efc %>% as_tibble())
```

```{r}
efc %>%
  frq(c161sex)
```
```{r}
efc %>% 
  group_by(e42dep) %>% 
  frq(c161sex)
```



```{r}
flat_table(efc, e42dep, c161sex)
```


```{r}
flat_table(efc, e42dep, c161sex, margin = "col")
```


```{r}
library(janitor)
efc %>% tabyl(e42dep, c161sex, show_na = FALSE)
```

## Coding


###  `rec()`

sjmisc

https://strengejacke.github.io/sjmisc/articles/recodingvariables.html

```{r}
efc$burden <- rec(
  efc$neg_c_7,
  rec = c("min:9=1 [low]; 10:12=2 [moderate]; 13:max=3 [high]; else=NA"),
  var.label = "Subjective burden",
  as.num = FALSE # we want a factor
)
# print frequencies
frq(efc$burden)
```

## Cutting


### `chop()` 

santoku

https://hughjonesd.github.io/santoku/tutorials/00-visualintroduction.html

```{r}
# devtools::install_github("hughjonesd/santoku")
# library(tidyverse) (load tidyverse before santoku to avoid conflicts)
library(santoku)
```

`cut()` in base R  

```{r}
x <- rnorm(100)
cut(x, 5) %>% table()  # 5 equal intervals
cut(x, -3:3) %>% table()
```

`ntile()` in dplyr:  

```{r}
ntile(x, 5) %>% table()
```




`chop()`

```{r}
chopped <- chop(x, breaks = -5:5)

chopped %>% table()

# chop() returns a factor
tibble(x, chopped)
```

If data is beyond the limits of `breaks`, they will be extended automatically, unless `extend = FALSE`, and values beyond the bounds will be converted to `NA`:  

```{r}
chopped <- chop(x, breaks = -1:1, extend = FALSE)
tibble(x, chopped)
```

To chop a single number into a separate category, put the number twice in `breaks`:

```{r}
x_zeros <- x 
x_zeros[1:5] <- 0

chopped <- chop(x_zeros, c(-1, 0, 0, 1))
tibble(x, chopped)
```

To quickly produce a table of chopped data, use `tab()`:  

```{r}
tab(x, breaks = -3:3)
```



